---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# wasmer: WebAssembly Runtime for R

<!-- badges: start -->
[![extendr](https://img.shields.io/badge/extendr-^0.8.1-276DC2)](https://extendr.github.io/extendr/extendr_api/)
<!-- badges: end -->

The `wasmer` package provides R bindings for the [Wasmer](https://wasmer.io/) WebAssembly runtime, allowing you to compile, instantiate, and execute WebAssembly modules directly from R. This opens up possibilities for high-performance computing, cross-language interoperability, and running untrusted code in a sandboxed environment.

## Basic Usage

### Initialize the Runtime

```{r init}
library(wasmer)

# Create the Wasmer runtime (must be called first)
runtime <- wasmer_runtime_new()
runtime
```

### Math Operations

```{r math-example}
# Test basic math operations
math_result <- wasmer_math_example_ext(runtime, 5L, 3L)
math_result

# Verify the results
stopifnot(math_result$add == 8)
stopifnot(math_result$multiply == 15)
```

## Advanced Usage

### Custom WebAssembly Modules

```{r fibonacci}
# Define a Fibonacci function in WebAssembly Text Format
fibonacci_wat <- '
(module
  (func $fibonacci (export "fibonacci") (param $n i32) (result i32)
    (if (i32.le_s (local.get $n) (i32.const 1))
      (then (return (local.get $n)))
    )
    (return (i32.add 
      (call $fibonacci (i32.sub (local.get $n) (i32.const 1)))
      (call $fibonacci (i32.sub (local.get $n) (i32.const 2)))
    ))
  )
)
'

# Compile the WebAssembly module
compile_result <- wasmer_compile_wat_ext(runtime, fibonacci_wat, "fibonacci_module")
compile_result

# Create an instance of the module
instance_result <- wasmer_instantiate_ext(runtime, "fibonacci_module", "fib_instance")
instance_result
# List available exports
exports <- wasmer_list_exports_ext(runtime, "fib_instance")
exports
stopifnot(exports$success == TRUE)
stopifnot("fibonacci" %in% exports$exports)
```

### Calculate Fibonacci Numbers

```{r fibonacci-test}
# Test fibonacci calculations with smaller numbers due to recursive implementation
fibonacci_tests <- data.frame(
  n = 0:8,
  expected = c(0, 1, 1, 2, 3, 5, 8, 13, 21)
)

fibonacci_tests$calculated <- sapply(fibonacci_tests$n, function(n) {
  result <- wasmer_call_function_ext(runtime, "fib_instance", "fibonacci", list(as.integer(n)))
  if (result$success) {
    return(result$values[[1]])
  } else {
    return(NA)
  }
})

fibonacci_tests

# Verify all calculations are correct
all_correct <- all(fibonacci_tests$calculated == fibonacci_tests$expected, na.rm = TRUE)
stopifnot(all_correct)
```

### Complex Example: Prime Number Checker

```{r prime-checker}
# Define a prime number checker in WebAssembly
prime_wat <- '
(module
  (func $is_prime (export "is_prime") (param $n i32) (result i32)
    (local $i i32)
    (local $sqrt_n i32)
    
    ;; Handle special cases
    (if (i32.lt_s (local.get $n) (i32.const 2))
      (then (return (i32.const 0))) ;; Not prime
    )
    
    (if (i32.eq (local.get $n) (i32.const 2))
      (then (return (i32.const 1))) ;; 2 is prime
    )
    
    ;; Check if even
    (if (i32.eqz (i32.rem_u (local.get $n) (i32.const 2)))
      (then (return (i32.const 0))) ;; Even numbers > 2 are not prime
    )
    
    ;; Check odd divisors up to sqrt(n)
    (local.set $i (i32.const 3))
    (local.set $sqrt_n (local.get $n)) ;; Approximate sqrt
    
    (loop $check_loop
      (if (i32.gt_s (i32.mul (local.get $i) (local.get $i)) (local.get $n))
        (then (return (i32.const 1))) ;; Prime
      )
      
      (if (i32.eqz (i32.rem_u (local.get $n) (local.get $i)))
        (then (return (i32.const 0))) ;; Not prime
      )
      
      (local.set $i (i32.add (local.get $i) (i32.const 2)))
      (br $check_loop)
    )
    
    (i32.const 1) ;; Should not reach here, but return prime
  )
)
'

# Compile and instantiate the prime checker
wasmer_compile_wat_ext(runtime, prime_wat, "prime_module")
wasmer_instantiate_ext(runtime, "prime_module", "prime_instance")

# Test with known primes and non-primes
test_numbers <- c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 13, 15, 17, 19, 21, 23, 25, 29, 31)
expected_primes <- c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29, 31)

results <- sapply(test_numbers, function(n) {
  result <- wasmer_call_function_ext(runtime, "prime_instance", "is_prime", list(as.integer(n)))
  return(result$values[[1]])
})

prime_results <- data.frame(
  number = test_numbers,
  is_prime = results == 1,
  expected = test_numbers %in% expected_primes
)

prime_results

# Verify all results are correct
all_prime_correct <- all(prime_results$is_prime == prime_results$expected)
stopifnot(all_prime_correct)
```

## Performance Comparison

```{r performance}
# Native R fibonacci implementation (recursive for fair comparison)
fib_r <- function(n) {
  if (n <= 1) return(n)
  return(fib_r(n-1) + fib_r(n-2))
}

# Benchmark both implementations
n_test <- 10L

bench_results <- bench::mark(
  wasm = wasmer_call_function_ext(runtime, "fib_instance", "fibonacci", list(n_test))$values[[1]],
  r = fib_r(n_test),
  check = FALSE,
  min_iterations = 5
)

bench_results
# Verify results match
stopifnot(bench_results$wasm[[1]] == bench_results$r[[1]])
```

## Running Tests

This README serves as both documentation and a test suite. To run the tests:

```{r tests, eval=FALSE}
# The chunks above serve as integration tests
# For additional unit tests, run:
tinytest::test_all(".")
```


## LLM Usage Disclosure
Code and documentation in this project have been generated with the assistance of the github Copilot LLM tools. While we have reviewed and edited the generated content, we acknowledge that LLM tools were used in the creation process and accordingly (since these models are trained on GPL code and other commons + proprietary software license is fake anyway) the code is released under GPL-3. So if you use this code in any way, you must comply with the GPL-3 license.