---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# wasmer: WebAssembly Runtime for R

<!-- badges: start -->
[![extendr](https://img.shields.io/badge/extendr-^0.8.1-276DC2)](https://extendr.github.io/extendr/extendr_api/)
<!-- badges: end -->

The `wasmer` package provides R bindings for the `rust` based [Wasmer](https://github.com/wasmerio/wasmer) WebAssembly runtime, allowing you to compile, instantiate, and execute WebAssembly modules directly from R. This opens up possibilities for high-performance computing, cross-language interoperability, and running untrusted code in a sandboxed environment. This is a wip.

## Installation
This package uses {rextenr}, so you will need a rust installation

```r
remotes::install_github("sounkou-bioinfo/wasmer")
```
##  Usage

### Initialize the Runtime

```{r init}
library(wasmer)

# Create the Wasmer runtime (must be called first)
runtime <- wasmer_runtime_new()
runtime
```

### Compiler Selection

You can choose the compiler backend when initializing the runtime. Supported compilers are `cranelift` (default, fast compilation) and `singlepass` (fastest compilation, linear time, good for blockchains/gas metering).

```{r init-compiler}
# Initialize with Cranelift (default)
rt_cranelift <- wasmer_runtime_new_with_compiler_ext("cranelift")

# Initialize with Singlepass
rt_singlepass <- wasmer_runtime_new_with_compiler_ext("singlepass")
```

### WASI Support

The package supports WASI (WebAssembly System Interface), allowing WASM modules to access system resources like files, environment variables, and clocks (if permitted).

```{r wasi}
# Create a runtime
rt_wasi <- wasmer_runtime_new()

# Create a WASI environment for a specific module name
# This must be done BEFORE instantiating the module
# It sets up stdout/stderr capture and other WASI imports
wasmer_wasi_state_new_ext(rt_wasi, "wasi_module")

# Compile and instantiate a module that uses WASI
# (Assuming you have a WASI-compliant WASM file or WAT)
# wasmer_compile_wasm_ext(rt_wasi, wasm_bytes, "wasi_module")
# wasmer_instantiate_ext(rt_wasi, "wasi_module", "wasi_instance")
```

### Math Operations compiled from Rust

```{r math-example}
# Test basic math operations
math_result <- wasmer_math_example_ext(runtime, 5L, 3L)
math_result

# Verify the results
stopifnot(math_result$add == 8)
stopifnot(math_result$multiply == 15)
```


### Custom WebAssembly Modules

```{r fibonacci}
# Define a Fibonacci function in WebAssembly Text Format
fibonacci_wat <- '
(module
  (func $fibonacci (export "fibonacci") (param $n i32) (result i32)
    (if (i32.le_s (local.get $n) (i32.const 1))
      (then (return (local.get $n)))
    )
    (return (i32.add 
      (call $fibonacci (i32.sub (local.get $n) (i32.const 1)))
      (call $fibonacci (i32.sub (local.get $n) (i32.const 2)))
    ))
  )
)
'

# Compile the WebAssembly module
compile_result <- wasmer_compile_wat_ext(runtime, fibonacci_wat, "fibonacci_module")
compile_result

# Create an instance of the module
instance_result <- wasmer_instantiate_ext(runtime, "fibonacci_module", "fib_instance")
instance_result
# List available exports
exports <- wasmer_list_exports_ext(runtime, "fib_instance")
exports
stopifnot(exports$success == TRUE)
stopifnot("fibonacci" %in% exports$exports)

# Native R fibonacci implementation 
fib_r <- function(n) {
  if (n <= 1) return(n)
  return(fib_r(n-1) + fib_r(n-2))
}
# Loop version of R



# Benchmark both implementations
n_test <- 10L

bench_results <- bench::mark(
  wasm = wasmer_call_function_ext(runtime, "fib_instance", "fibonacci", list(n_test))$values[[1]],
  r = fib_r(n_test),
  check = FALSE,
  min_iterations = 5
)

bench_results
stopifnot(bench_results$wasm[[1]] == bench_results$r[[1]])
```

### Calculate Fibonacci Numbers

```{r fibonacci-test}
# Test fibonacci calculations with smaller numbers due to recursive implementation
fibonacci_tests <- data.frame(
  n = 0:8,
  expected = c(0, 1, 1, 2, 3, 5, 8, 13, 21)
)

fibonacci_tests$calculated <- sapply(fibonacci_tests$n, function(n) {
  result <- wasmer_call_function_ext(runtime, "fib_instance", "fibonacci", list(as.integer(n)))
  if (result$success) {
    return(result$values[[1]])
  } else {
    return(NA)
  }
})

fibonacci_tests

# Verify all calculations are correct
all_correct <- all(fibonacci_tests$calculated == fibonacci_tests$expected, na.rm = TRUE)
stopifnot(all_correct)
```

###  Prime Number Checker

```{r prime-checker}
# Define a prime number checker in WebAssembly
prime_wat <- '
(module
  (func $is_prime (export "is_prime") (param $n i32) (result i32)
    (local $i i32)
    (local $sqrt_n i32)
    
    ;; Handle special cases
    (if (i32.lt_s (local.get $n) (i32.const 2))
      (then (return (i32.const 0))) ;; Not prime
    )
    
    (if (i32.eq (local.get $n) (i32.const 2))
      (then (return (i32.const 1))) ;; 2 is prime
    )
    
    ;; Check if even
    (if (i32.eqz (i32.rem_u (local.get $n) (i32.const 2)))
      (then (return (i32.const 0))) ;; Even numbers > 2 are not prime
    )
    
    ;; Check odd divisors up to sqrt(n)
    (local.set $i (i32.const 3))
    (local.set $sqrt_n (local.get $n)) ;; Approximate sqrt
    
    (loop $check_loop
      (if (i32.gt_s (i32.mul (local.get $i) (local.get $i)) (local.get $n))
        (then (return (i32.const 1))) ;; Prime
      )
      
      (if (i32.eqz (i32.rem_u (local.get $n) (local.get $i)))
        (then (return (i32.const 0))) ;; Not prime
      )
      
      (local.set $i (i32.add (local.get $i) (i32.const 2)))
      (br $check_loop)
    )
    
    (i32.const 1) ;; Should not reach here, but return prime
  )
)
'

# Compile and instantiate the prime checker
wasmer_compile_wat_ext(runtime, prime_wat, "prime_module")
wasmer_instantiate_ext(runtime, "prime_module", "prime_instance")

# Test with known primes and non-primes
test_numbers <- c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 13, 15, 17, 19, 21, 23, 25, 29, 31)
expected_primes <- c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29, 31)

results <- sapply(test_numbers, function(n) {
  result <- wasmer_call_function_ext(runtime, "prime_instance", "is_prime", list(as.integer(n)))
  return(result$values[[1]])
})

prime_results <- data.frame(
  number = test_numbers,
  is_prime = results == 1,
  expected = test_numbers %in% expected_primes
)

prime_results

# Verify all results are correct
all_prime_correct <- all(prime_results$is_prime == prime_results$expected)
stopifnot(all_prime_correct)
```


## Inspect Exported Function Signatures

```{r exported-signatures}
# After instantiating a module, you can inspect its exported function signatures
signatures <- wasmer_list_function_signatures_ext(runtime, "fib_instance")
print(signatures)

# You can also inspect other instances
signatures_prime <- wasmer_list_function_signatures_ext(runtime, "prime_instance")
print(signatures_prime)
```

## WAT to WASM Binary Conversion and Binary Module Loading

```{r wat-to-wasm-binary-roundtrip}
# Convert WAT to WASM binary
wat_code <- ' (module (func $double (export "double") (param $x i32) (result i32) local.get $x i32.const 2 i32.mul) ) '
wasm_bin <- wasmer_wat_to_wasm_ext(wat_code)
stopifnot(is.raw(wasm_bin))
stopifnot(length(wasm_bin) > 0)

# Compile the binary WASM into a module
compile_bin_result <- wasmer_compile_wasm_ext(runtime, wasm_bin, "double_module_bin")
stopifnot(is.character(compile_bin_result))
stopifnot(grepl("compiled from binary successfully", compile_bin_result, ignore.case = TRUE))

# Instantiate and call the exported function
instance_bin_result <- wasmer_instantiate_ext(runtime, "double_module_bin", "double_instance_bin")
stopifnot(is.character(instance_bin_result))
stopifnot(grepl("created successfully", instance_bin_result, ignore.case = TRUE))

# Call the function
result_bin <- wasmer_call_function_ext(runtime, "double_instance_bin", "double", list(21L))
stopifnot(result_bin$success)
stopifnot(result_bin$values[[1]] == 42)
```


## Register and Call an R Function from WASM

```{r r-host-function-example}
# Define an R function to double a value
r_double <- function(x) as.integer(x * 2)
wasmer_register_r_function_ext(runtime, "r_double", r_double)

wat_code <- '
(module
  (import "env" "r_host_call" (func $r_host_call (param i32 i32 i32 i32) (result i32)))
  (memory (export "memory") 1)
  (data (i32.const 0) "r_double")
  (func $call_r_double (export "call_r_double") (param $x i32) (result i32)
    (i32.store (i32.const 100) (local.get $x))
    (call $r_host_call (i32.const 0) (i32.const 8) (i32.const 1) (i32.const 100))
  )
)
'

compile_result <- wasmer_compile_wat_ext(runtime, wat_code, "rhost_module")
compile_result
result <- wasmer_call_function_ext(runtime, "rhost_instance", "call_r_double", list(21L))
stopifnot(result$values[[1]] == 42)
```


## Simple WASM Memory Example

This example demonstrates how to use Wasmer's memory utilities from R.

```{r memory}
# Create a runtime
rt <- wasmer_runtime_new()

# Compile a simple WAT module with exported memory
wat <- '(module (memory (export "memory") 1) (func (export "write") (param i32 i32) (result i32)
  local.get 0
  local.get 1
  i32.store
  local.get 0))'
wasmer_compile_wat_ext(rt, wat, "memmod")
wasmer_instantiate_ext(rt, "memmod", "inst")

# Write bytes to memory
wasmer_memory_write_ext(rt, "inst", "memory", 0, as.raw(c(65, 66, 67))) # Write 'A', 'B', 'C' at offset 0

# Read bytes from memory
bytes <- wasmer_memory_read_ext(rt, "inst", "memory", 0, 3)
print(bytes) # Should print raw vector: 41 42 43

# Grow memory by 1 page (64KiB)
success <- wasmer_memory_grow_ext(rt, "inst", "memory", 1L)
print(success) # Should print TRUE if memory was grown

# Read as string
str <- wasmer_memory_read_string_ext(rt, "inst", "memory", 0, 3)
print(str) # Should print "ABC"
```

This example shows how to write and read bytes and strings from WASM memory using Wasmer's R interface.


## Generic WASM Table and Dynamic Host Function Example

This example demonstrates two approaches:
- How to use a WASM table with static host functions (required for table/funcref)
- How to use dynamic host functions for direct import/export (not tables)

```{r tables}
# Create runtime
rt <- wasmer_runtime_new()

# --- Table Example (static host function) ---
# Note: Static host functions are generated and exported via macros in Rust for common signatures.
#       Only these can be inserted into WASM tables (funcref) due to Wasmer backend restrictions.
wat_table <- ' (module
  (type $callback_t (func (param i32 i32) (result i32)))
  (type $call_callback_t (func (param i32 i32 i32) (result i32)))
  (table $t1 3 6 funcref)
  (func $default_fn (type $callback_t) (param $a i32) (param $b i32) (result i32)
     (i32.add (i32.mul (local.get $a) (i32.const 2)) (i32.mul (local.get $b) (i32.const 2))))
  (elem $t1 (i32.const 0) $default_fn $default_fn $default_fn)
  (func $call_callback (type $call_callback_t) (param $idx i32) (param $arg1 i32) (param $arg2 i32) (result i32)
    (call_indirect (type $callback_t) (local.get $arg1) (local.get $arg2) (local.get $idx)))
  (export "call_callback" (func $call_callback))
  (export "__indirect_function_table" (table $t1))
)'

wasmer_compile_wat_ext(rt, wat_table, "mod_table")
wasmer_instantiate_ext(rt, "mod_table", "inst_table")

# Create a WASM table from R
r_table_ptr <- wasmer_table_new_ext(rt, 3L, 6L)

# Create a host function (sum) with static signature (i32, i32) -> i32
# This uses a macro-generated wrapper in Rust (see package source for supported signatures)
host_sum <- function(x, y) as.integer(x + y)
host_func <- wasmer_function_new_static_ext(rt, host_sum, "host_sum")

# Set table index 1 to host function (SUPPORTED with static signature)
wasmer_table_set_ext(rt, r_table_ptr, 1L, host_func)

# Call the WASM function via table (note: this will not affect the instance's table)
result <- wasmer_call_function_ext(rt, "inst_table", "call_callback", list(1L, 2L, 7L))
result$values[[1]] # Will print 18 (default_fn)

# --- Dynamic Host Function Example (direct import, not table) ---
wat_dyn <- ' (module
  (import "env" "host_sum_dyn" (func $host_sum_dyn (param i32 i32) (result i32)))
  (func (export "call_host_sum_dyn") (param $a i32) (param $b i32) (result i32)
    (call $host_sum_dyn (local.get $a) (local.get $b))
  )
)'

wasmer_compile_wat_ext(rt, wat_dyn, "mod_dyn")

# Create a dynamic host function (can be any signature)
host_sum_dyn <- function(x, y) as.integer(x + y)
host_func_dyn <- wasmer_function_new_ext(rt, host_sum_dyn, c("i32", "i32"), c("i32"), "host_sum_dyn")

# Note: You must wire up the import manually when instantiating (see package docs)
# For this example, the import is handled automatically if the registry name matches
wasmer_instantiate_ext(rt, "mod_dyn", "inst_dyn")

# Call the WASM function that uses the dynamic host function
result_dyn <- wasmer_call_function_ext(rt, "inst_dyn", "call_host_sum_dyn", list(2L, 7L))
result_dyn$values[[1]] # Should print 9

```

> [!IMPORTANT]
> **Dynamic vs Static Host Functions & Tables**
> 
> There is a known limitation ("pending issue") regarding **Dynamic Host Functions** (`wasmer_function_new_ext`) and WASM Tables (`funcref`).
> 
> - **Dynamic Host Functions**: Created with `wasmer_function_new_ext`. These are flexible and can have any signature, but they **CANNOT** be inserted into WASM Tables or used as `funcref`. Attempting to do so will fail or cause runtime errors because they lack the necessary typed trampoline at the engine level.
> - **Static Host Functions**: Created with `wasmer_function_new_static_ext` (or typed variants like `wasmer_function_new_i32_to_i32`). These have fixed signatures known at compile time. These **CAN** be inserted into Tables and used as `funcref`.
>
> If you need to use host functions in a Table (e.g., for indirect calls), you **MUST** use the static variants.


## LLM Usage Disclosure
Code and documentation in this project have been generated with the assistance of the github Copilot LLM tools. While we have reviewed and edited the generated content, we acknowledge that LLM tools were used in the creation process and accordingly (since these models are trained on GPL code and other commons + proprietary software license is fake anyway) the code is released under GPL-3. So if you use this code in any way, you must comply with the GPL-3 license.