---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# wasmer: WebAssembly Runtime for R

<!-- badges: start -->
[![extendr](https://img.shields.io/badge/extendr-^0.8.1-276DC2)](https://extendr.github.io/extendr/extendr_api/)
<!-- badges: end -->

The `wasmer` package provides R bindings for the `rust` based [Wasmer](https://github.com/wasmerio/wasmer) WebAssembly runtime, allowing you to compile, instantiate, and execute WebAssembly modules directly from R. This opens up possibilities for high-performance computing, cross-language interoperability, and running untrusted code in a sandboxed environment. This is a wip and may be dropped for a pure C runtime.

## Installation
This package uses [{rextendr}](https://github.com/extendr/rextendr), so you will need a Rust installation

```r
remotes::install_github("sounkou-bioinfo/wasmer")
or
install.packages("wasmer", repos = c('https://sounkou-bioinfo.r-universe.dev', 'https://cloud.r-project.org'))
```
##  Usage

### Initialize the Runtime with specified compiler


You can choose the compiler backend when initializing the runtime. Each compiler offers different trade-offs:

- **Cranelift** (default): Fast compilation with good runtime performance
- **Singlepass**: Fastest compilation (linear time), ideal for blockchains/gas metering
- **LLVM** (optional): Best runtime performance (~50% faster than Cranelift), but requires LLVM 18 installed

```{r init}
library(wasmer)

# Create the Wasmer runtime (must be called first)
runtime <- wasmer_runtime_new()
runtime
# can release ressources before the gc and the external pointer release
# mechamism
wasmer_runtime_release_ressources(runtime)
rm(runtime)
# Initialize with Cranelift (default)
runtime <- wasmer_runtime_new_with_compiler_ext("cranelift")
rm(runtime)
# Initialize with Singlepass for fastest compilation
runtime <- wasmer_runtime_new_with_compiler_ext("singlepass")
rm(runtime)
gc()
# LLVM is available only if package was built with LLVM support
# (requires LLVM 18 on system)
#runtime <- wasmer_runtime_new_with_compiler_ext("llvm")
# to continue with the toor, we create a runtime that will be re-using
# only one runtime per process is recommended
runtime <- wasmer_runtime_new()
runtime
```

### Compiler Selection


```{r init-compiler}


```



### Math Operations compiled from Rust 

This is to test basic host functionality using functions in rust.

```{r math-example}
# Test basic math operations
math_result <- wasmer_math_example_ext(runtime, 5L, 3L)
math_result

# Verify the results
stopifnot(math_result$add == 8)
stopifnot(math_result$multiply == 15)
```


### Custom WebAssembly Modules

Basic module tests

#### Fibonaci

```{r fibonacci}
# Define a Fibonacci function in WebAssembly Text Format
fibonacci_wat <- '
(module
  (func $fibonacci (export "fibonacci") (param $n i32) (result i32)
    (if (i32.le_s (local.get $n) (i32.const 1))
      (then (return (local.get $n)))
    )
    (return (i32.add 
      (call $fibonacci (i32.sub (local.get $n) (i32.const 1)))
      (call $fibonacci (i32.sub (local.get $n) (i32.const 2)))
    ))
  )
)
'

# Compile the WebAssembly module
compile_result <- wasmer_compile_wat_ext(runtime, fibonacci_wat, "fibonacci_module")
compile_result

# Create an instance of the module
instance_result <- wasmer_instantiate_ext(runtime, "fibonacci_module", "fib_instance")
instance_result
# List available exports
exports <- wasmer_list_exports_ext(runtime, "fib_instance")
exports
stopifnot(exports$success == TRUE)
stopifnot("fibonacci" %in% exports$exports)

# Native R fibonacci implementations
# Naive recursive (grows stack)
fib_r <- function(n) {
  if (n <= 1) return(n)
  return(fib_r(n-1) + fib_r(n-2))
}

# Tail-recursive using Tailcall 
fib_tailcall <- function(n) {
  fib_iter <- function(a, b, count) {
    if (count == 0) {
      a
    } else {
      # Force evaluation to avoid deferred computations
      new_a <- b
      new_b <- a + b
      Tailcall(fib_iter, new_a, new_b, count - 1)
    }
  }
  fib_iter(0L, 1L, n)
}

# Benchmark all three implementations
n_test <- 20L

bench_results <- bench::mark(
  wasm = wasmer_call_function_ext(runtime, "fib_instance", "fibonacci", list(n_test))$values[[1]],
  r_naive = fib_r(n_test),
  r_tailcall = fib_tailcall(n_test),
  check = FALSE,
  min_iterations = 10
)

bench_results
stopifnot(bench_results$wasm[[1]] == bench_results$r_naive[[1]])
stopifnot(bench_results$r_naive[[1]] == bench_results$r_tailcall[[1]])
```

#### Calculate Fibonacci Numbers

```{r fibonacci-test}
# Test fibonacci calculations with smaller numbers due to recursive implementation
fibonacci_tests <- data.frame(
  n = 0:8,
  expected = c(0, 1, 1, 2, 3, 5, 8, 13, 21)
)

fibonacci_tests$calculated <- sapply(fibonacci_tests$n, function(n) {
  result <- wasmer_call_function_ext(runtime, "fib_instance", "fibonacci", list(as.integer(n)))
  if (result$success) {
    return(result$values[[1]])
  } else {
    return(NA)
  }
})

fibonacci_tests

# Verify all calculations are correct
all_correct <- all(fibonacci_tests$calculated == fibonacci_tests$expected, na.rm = TRUE)
stopifnot(all_correct)
```

###  Prime Number Checker

```{r prime-checker}
# Define a prime number checker in WebAssembly
prime_wat <- '
(module
  (func $is_prime (export "is_prime") (param $n i32) (result i32)
    (local $i i32)
    (local $sqrt_n i32)
    
    ;; Handle special cases
    (if (i32.lt_s (local.get $n) (i32.const 2))
      (then (return (i32.const 0))) ;; Not prime
    )
    
    (if (i32.eq (local.get $n) (i32.const 2))
      (then (return (i32.const 1))) ;; 2 is prime
    )
    
    ;; Check if even
    (if (i32.eqz (i32.rem_u (local.get $n) (i32.const 2)))
      (then (return (i32.const 0))) ;; Even numbers > 2 are not prime
    )
    
    ;; Check odd divisors up to sqrt(n)
    (local.set $i (i32.const 3))
    (local.set $sqrt_n (local.get $n)) ;; Approximate sqrt
    
    (loop $check_loop
      (if (i32.gt_s (i32.mul (local.get $i) (local.get $i)) (local.get $n))
        (then (return (i32.const 1))) ;; Prime
      )
      
      (if (i32.eqz (i32.rem_u (local.get $n) (local.get $i)))
        (then (return (i32.const 0))) ;; Not prime
      )
      
      (local.set $i (i32.add (local.get $i) (i32.const 2)))
      (br $check_loop)
    )
    
    (i32.const 1) ;; Should not reach here, but return prime
  )
)
'

# Compile and instantiate the prime checker
wasmer_compile_wat_ext(runtime, prime_wat, "prime_module")
wasmer_instantiate_ext(runtime, "prime_module", "prime_instance")

# Test with known primes and non-primes
test_numbers <- c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 13, 15, 17, 19, 21, 23, 25, 29, 31)
expected_primes <- c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29, 31)

results <- sapply(test_numbers, function(n) {
  result <- wasmer_call_function_ext(runtime, "prime_instance", "is_prime", list(as.integer(n)))
  return(result$values[[1]])
})

prime_results <- data.frame(
  number = test_numbers,
  is_prime = results == 1,
  expected = test_numbers %in% expected_primes
)

prime_results

# Verify all results are correct
all_prime_correct <- all(prime_results$is_prime == prime_results$expected)
stopifnot(all_prime_correct)
```


## Inspect Exported Function Signatures

```{r exported-signatures}
# After instantiating a module, you can inspect its exported function signatures
signatures <- wasmer_list_function_signatures_ext(runtime, "fib_instance")
signatures

# You can also inspect other instances
signatures_prime <- wasmer_list_function_signatures_ext(runtime, "prime_instance")
signatures_prime
```

## WAT to WASM Binary Conversion and Binary Module Loading

```{r wat-to-wasm-binary-roundtrip}
# Convert WAT to WASM binary
wat_code <- ' (module (func $double (export "double") (param $x i32) (result i32) local.get $x i32.const 2 i32.mul) ) '
wasm_bin <- wasmer_wat_to_wasm_ext(wat_code)
stopifnot(is.raw(wasm_bin))
stopifnot(length(wasm_bin) > 0)

# Compile the binary WASM into a module
compile_bin_result <- wasmer_compile_wasm_ext(runtime, wasm_bin, "double_module_bin")
stopifnot(is.character(compile_bin_result))
stopifnot(grepl("compiled from binary successfully", compile_bin_result, ignore.case = TRUE))

# Instantiate and call the exported function
instance_bin_result <- wasmer_instantiate_ext(runtime, "double_module_bin", "double_instance_bin")
stopifnot(is.character(instance_bin_result))
stopifnot(grepl("created successfully", instance_bin_result, ignore.case = TRUE))

# Call the function
result_bin <- wasmer_call_function_ext(runtime, "double_instance_bin", "double", list(21L))
stopifnot(result_bin$success)
stopifnot(result_bin$values[[1]] == 42)
result_bin$values[[1]]
```


## Register and Call an R Function from WASM
This interface is very limited, refer to the function tables example.
```{r r-host-function-example}


## Corrected Example: Register and Call an R Function from WASM

# Define an R function to double a value
r_double <- function(a) {
  2L * as.integer(a)
}
handle <- wasmer_register_r_function_ext(runtime, "r_double", r_double)

# The Rust host function expects (handle, args_ptr, argc), but only the args are passed to R
wat_code <- sprintf('
(module
  (import "env" "r_host_call" (func $r_host_call (param i32 i32 i32) (result i32)))
  (memory (export "memory") 1)
  (data (i32.const 0) "r_double")
  (func $call_r_double (export "call_r_double") (param $x i32) (result i32)
    (i32.store (i32.const 100) (local.get $x))
    (call $r_host_call (i32.const %d) (i32.const 100) (i32.const 1))
  )
)
', handle)

compile_result <- wasmer_compile_wat_ext(runtime, wat_code, "rhost_module")
compile_result
instance_result <- wasmer_instantiate_ext(runtime, "rhost_module", "rhost_instance")
instance_result
result <- wasmer_call_function_ext(runtime, "rhost_instance", "call_r_double", list(21L))
result
stopifnot(result$values[[1]] == 42)
```

## WASM Memory

### Simple WASM Memory Example

This example shows how to write and read bytes and strings from WASM memory using Wasmer's R interface.

```{r memory}
# Compile a simple WAT module with exported memory
wat <- '(module (memory (export "memory") 1) (func (export "write") (param i32 i32) (result i32)
  local.get 0
  local.get 1
  i32.store
  local.get 0))'
wasmer_compile_wat_ext(runtime, wat, "memmod")
wasmer_instantiate_ext(runtime, "memmod", "inst")

# Write bytes to memory
wasmer_memory_write_ext(runtime, "inst", "memory", 0, as.raw(c(65, 66, 67))) # Write 'A', 'B', 'C' at offset 0

# Read bytes from memory
bytes <- wasmer_memory_read_ext(runtime, "inst", "memory", 0, 3)
bytes # Should print raw vector: 41 42 43

# Grow memory by 1 page (64KiB)
success <- wasmer_memory_grow_ext(runtime, "inst", "memory", 1L)
success # Should print TRUE if memory was grown

# Read as string
str <- wasmer_memory_read_string_ext(runtime, "inst", "memory", 0, 3)
str # Should print "ABC"
```

### Write R Array, Sum in WASM

This example demonstrates passing a numeric array from R to WASM memory, summing it in WASM, and reading the result back.

```{r memory-demo}
# WASM module that sums an array of f64 values in memory
wat_sum_array <- '
(module
  (memory (export "memory") 1)
  (func $sum_array (export "sum_array") (param $offset i32) (param $len i32) (result f64)
    (local $i i32)
    (local $sum f64)
    (local.set $i (i32.const 0))
    (local.set $sum (f64.const 0))
    (block $break
      (loop $loop
        (br_if $break (i32.ge_u (local.get $i) (local.get $len)))
        ;; Load f64 from memory: memory is byte-addressed, f64 is 8 bytes
        (local.set $sum
          (f64.add
            (local.get $sum)
            (f64.load (i32.add (local.get $offset) (i32.mul (local.get $i) (i32.const 8))))
          )
        )
        (local.set $i (i32.add (local.get $i) (i32.const 1)))
        (br $loop)
      )
    )
    (local.get $sum)
  )
)
'

# Compile and instantiate
wasmer_compile_wat_ext(runtime, wat_sum_array, "sum_mod")
wasmer_instantiate_ext(runtime, "sum_mod", "sum_inst")

# Generate R array and write to WASM memory
arr <- rnorm(10)
offset <- 0L  # start of memory
raw_bytes <- writeBin(arr, raw(), size=8)  # f64 is 8 bytes
wasmer_memory_write_ext(runtime, "sum_inst", "memory", offset, raw_bytes)

# Call WASM sum_array(offset, length)
result <- wasmer_call_function_ext(runtime, "sum_inst", "sum_array", list(offset, length(arr)))
result
sum(arr)
stopifnot(abs(sum(arr) - result$values[[1]]) < 1e-8)
```


## WASM Tables and Typed Host Functions

This example demonstrates using typed host functions that can be inserted into WASM tables.

```{r tables}
# Create runtime
rt <- runtime

# Example 1: (i32, i32) -> i32 - Binary operations
host_add <- function(x, y) as.integer(x + y)
host_multiply <- function(x, y) as.integer(x * y)

add_func <- wasmer_function_new_i32_i32_to_i32(runtime, host_add)
mul_func <- wasmer_function_new_i32_i32_to_i32(runtime, host_multiply)

# Create and use in WASM table
table_ptr <- wasmer_table_new_ext(runtime, 2L, 10L)
wasmer_table_set_ext(runtime, table_ptr, 0L, add_func)
wasmer_table_set_ext(runtime, table_ptr, 1L, mul_func)

# Example 2: (i32) -> i32 - Unary integer operations
host_square <- function(x) as.integer(x * x)
host_double <- function(x) as.integer(x * 2)

square_func <- wasmer_function_new_i32_to_i32(runtime, host_square)
double_func <- wasmer_function_new_i32_to_i32(runtime, host_double)

# Example 3: (f64, f64) -> f64 - Binary float operations  
host_avg <- function(x, y) (x + y) / 2
host_max <- function(x, y) max(x, y)

avg_func <- wasmer_function_new_f64_f64_to_f64(runtime, host_avg)
max_func <- wasmer_function_new_f64_f64_to_f64(runtime, host_max)

# Example 4: (f64) -> f64 - Unary float operations
host_sqrt <- function(x) sqrt(x)
host_abs <- function(x) abs(x)

sqrt_func <- wasmer_function_new_f64_to_f64(runtime, host_sqrt)
abs_func <- wasmer_function_new_f64_to_f64(runtime, host_abs)

# Example 5: (i32) -> void - Side effects (logging/printing)
logged_values <- c()
host_log <- function(x) {
  logged_values <<- c(logged_values, x)
  invisible(NULL)
}

log_func <- wasmer_function_new_i32_to_void(runtime, host_log)

# Example 6: () -> i32 - Generators/timestamps
counter <- 0
host_next_id <- function() {
  counter <<- counter + 1
  as.integer(counter)
}

next_id_func <- wasmer_function_new_void_to_i32(runtime, host_next_id)

# Now create simple WASM modules that actually CALL these host functions!
## Example: Call Host Function from WASM Table


# Reuse the same runtime used above
# Host function: next_id generator
counter <- 0
host_next_id <- function() {
  counter <<- counter + 1
  as.integer(counter)
}
next_id_func <- wasmer_function_new_void_to_i32(runtime, host_next_id)

# Create WASM table and set host function

# Create WASM table with 4 slots and set host functions

# Create WASM table with 5 slots and set host functions

# Create WASM table with 9 slots and set host functions
table_ptr <- wasmer_table_new_ext(runtime, 9L, 9L)
wasmer_table_set_ext(runtime, table_ptr, 0L, next_id_func)   # slot 0: next_id
wasmer_table_set_ext(runtime, table_ptr, 1L, add_func)       # slot 1: add
wasmer_table_set_ext(runtime, table_ptr, 2L, square_func)    # slot 2: square
wasmer_table_set_ext(runtime, table_ptr, 3L, mul_func)       # slot 3: multiply
wasmer_table_set_ext(runtime, table_ptr, 4L, double_func)    # slot 4: double
wasmer_table_set_ext(runtime, table_ptr, 5L, avg_func)       # slot 5: avg (f64, f64) -> f64
wasmer_table_set_ext(runtime, table_ptr, 6L, max_func)       # slot 6: max (f64, f64) -> f64
wasmer_table_set_ext(runtime, table_ptr, 7L, sqrt_func)      # slot 7: sqrt (f64) -> f64
wasmer_table_set_ext(runtime, table_ptr, 8L, abs_func)       # slot 8: abs (f64) -> f64

# WASM module that imports and uses a table
wat_table_call <- '
(module
  ;; Types for each host function signature
  (type $gen (func (result i32)))
  (type $binop (func (param i32 i32) (result i32)))
  (type $unop (func (param i32) (result i32)))
  (type $binopf64 (func (param f64 f64) (result f64)))
  (type $unopf64 (func (param f64) (result f64)))

  ;; Import a table with 9 funcref slots
  (import "env" "host_table" (table 9 funcref))

  ;; Call next_id (slot 0)
  (func $call_next_id (export "call_next_id") (result i32)
    (call_indirect (type $gen) (i32.const 0))
  )
  ;; Call add (slot 1)
  (func $call_add (export "call_add") (param $x i32) (param $y i32) (result i32)
    (call_indirect (type $binop) (local.get $x) (local.get $y) (i32.const 1))
  )
  ;; Call square (slot 2)
  (func $call_square (export "call_square") (param $x i32) (result i32)
    (call_indirect (type $unop) (local.get $x) (i32.const 2))
  )
  ;; Call multiply (slot 3)
  (func $call_multiply (export "call_multiply") (param $x i32) (param $y i32) (result i32)
    (call_indirect (type $binop) (local.get $x) (local.get $y) (i32.const 3))
  )
  ;; Call double (slot 4)
  (func $call_double (export "call_double") (param $x i32) (result i32)
    (call_indirect (type $unop) (local.get $x) (i32.const 4))
  )
  ;; Call avg (slot 5)
  (func $call_avg (export "call_avg") (param $x f64) (param $y f64) (result f64)
    (call_indirect (type $binopf64) (local.get $x) (local.get $y) (i32.const 5))
  )
  ;; Call max (slot 6)
  (func $call_max (export "call_max") (param $x f64) (param $y f64) (result f64)
    (call_indirect (type $binopf64) (local.get $x) (local.get $y) (i32.const 6))
  )
  ;; Call sqrt (slot 7)
  (func $call_sqrt (export "call_sqrt") (param $x f64) (result f64)
    (call_indirect (type $unopf64) (local.get $x) (i32.const 7))
  )
  ;; Call abs (slot 8)
  (func $call_abs (export "call_abs") (param $x f64) (result f64)
    (call_indirect (type $unopf64) (local.get $x) (i32.const 8))
  )
)
'

# Compile the module
wasmer_compile_wat_ext(runtime, wat_table_call, "table_module")

# Instantiate the module, passing the table as an import
wasmer_instantiate_with_table_ext(runtime, "table_module", "table_instance", table_ptr)

# Call the WASM function, which calls the host function via the table

# Call each WASM function, which calls the corresponding host function via the table
result_next_id <- wasmer_call_function_ext(runtime, "table_instance", "call_next_id", list())
result_next_id

result_add <- wasmer_call_function_ext(runtime, "table_instance", "call_add", list(10L, 32L))
result_add

result_square <- wasmer_call_function_ext(runtime, "table_instance", "call_square", list(7L))
result_square

result_multiply <- wasmer_call_function_ext(runtime, "table_instance", "call_multiply", list(6L, 7L))
result_multiply

result_double <- wasmer_call_function_ext(runtime, "table_instance", "call_double", list(21L))
result_double

result_avg <- wasmer_call_function_ext(runtime, "table_instance", "call_avg", list(1.5, 2.5))
result_avg

result_max <- wasmer_call_function_ext(runtime, "table_instance", "call_max", list(1.5, 2.5))
result_max

result_sqrt <- wasmer_call_function_ext(runtime, "table_instance", "call_sqrt", list(9.0))
result_sqrt
result_sqrt <- wasmer_call_function_ext(runtime, "table_instance", "call_sqrt", list(9.115))
result_sqrt
result_abs <- wasmer_call_function_ext(runtime, "table_instance", "call_abs", list(-42.0))
result_abs
```

### Available Typed Host Function Signatures

The package provides the following typed host function creators:

- `wasmer_function_new_i32_to_i32(runtime, rfun)` - (i32) -> i32
- `wasmer_function_new_i32_i32_to_i32(runtime, rfun)` - (i32, i32) -> i32  
- `wasmer_function_new_f64_f64_to_f64(runtime, rfun)` - (f64, f64) -> f64
- `wasmer_function_new_f64_to_f64(runtime, rfun)` - (f64) -> f64
- `wasmer_function_new_i32_to_void(runtime, rfun)` - (i32) -> void (for logging)
- `wasmer_function_new_void_to_i32(runtime, rfun)` - () -> i32 (for generators)


## WASI(X)

The package supports WASIX (WebAssembly System Interface Extension), allowing WASM modules to interact with system resources and files in a sandboxed way.

```{r wasix}
# Create a WASIX-enabled runtime
rt_wasix <- wasmer_runtime_new()

# Create a temp directory and file in R
output_dir <- tempdir()
output_file <- file.path(output_dir, "wasix_output.txt")
file.create(output_file)
file_name <- basename(output_file)
file_data <- "Hello from WASIX file"

# Create WASIX environment 
wasmer_wasi_state_new_ext(rt_wasix, "hello_wasix", "wasix")

# WASIX WAT: open file and write to it using path_open and fd_write
escape_wat_string <- function(x) {
  paste0(gsub('([\\"])', '\\\\1', x), collapse = "")
}
file_wasix_wat <- sprintf('
(module
  (import "wasix_32v1" "path_open"
    (func $path_open (param i32 i32 i32 i32 i32 i64 i64 i32 i32) (result i32)))
  (import "wasix_32v1" "fd_write"
    (func $fd_write (param i32 i32 i32 i32) (result i32)))
  (memory (export "memory") 1)
  (data (i32.const 0) "%s")
  (data (i32.const 100) "%s")
  (func $_start (export "_start")
    (local $fd i32)
    ;; Open the file in the preopened dir (fd=3)
    (i32.store (i32.const 200) (i32.const 0))   ;; pointer to filename
    (i32.store (i32.const 204) (i32.const %d))  ;; length of filename
    (drop (call $path_open
      (i32.const 3)   ;; preopened dir fd
      (i32.const 0)   ;; dirflags
      (i32.const 0)   ;; path_ptr
      (i32.const %d)  ;; path_len
      (i32.const 1)   ;; oflags (create)
      (i64.const 0x0000000000000002) ;; rights_base (RIGHT_FD_WRITE)
      (i64.const 0x0000000000000000) ;; rights_inheriting
      (i32.const 0)   ;; fdflags
      (i32.const 208) ;; fd_out_ptr
    ))
    ;; fd is now at memory[208]
    (local.set $fd (i32.load (i32.const 208)))
    ;; Prepare iovec for writing
    (i32.store (i32.const 300) (i32.const 100)) ;; ptr to string
    (i32.store (i32.const 304) (i32.const %d))  ;; length
    ;; Write to file
    (drop (call $fd_write
      (local.get $fd)
      (i32.const 300)
      (i32.const 1)
      (i32.const 400)
    ))
  )
)
', escape_wat_string(file_name), escape_wat_string(file_data), nchar(file_name), nchar(file_name), nchar(file_data))

# Compile and instantiate the module
wasmer_compile_wat_ext(rt_wasix, file_wasix_wat, "hello_wasix")
wasmer_instantiate_ext(rt_wasix, "hello_wasix", "file_instance")

# Call the WASIX _start function (writes to file)
result <- wasmer_call_function_ext(rt_wasix, "file_instance", "_start", list())
result

# Read the output file in R
readLines(output_file)
```
## Pending issues


- [ ] WASI exploration

- [ ] Compiler and engine interface

- [ ] Tests, tests, tests



## LLM Usage Disclosure
Code and documentation in this project have been generated with the assistance of the github Copilot LLM tools. While we have reviewed and edited the generated content, we acknowledge that LLM tools were used in the creation process and accordingly (since these models are trained on GPL code and other commons + proprietary software license is fake anyway) the code is released under GPL-3. So if you use this code in any way, you must comply with the GPL-3 license.